---
title: "molecool_Module9"
author: "molecool"
date: "12/10/2021"
output: html_document
bibliography: BIOL3140.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(tidyverse)
library(ape)
library(nlme)
library(geiger)
library(caper)
library(phytools)
library(viridis)
library(MuMIn)
library(rgbif)
library(MuMIn)
library(rnoaa)
library(ggmap)
library(usmap)
library(magick)
library(cowplot)
library(lme4) 
library(car) 
library(data.table) 
library(dplyr)
library(knitr)
```
## Introduction

Who has never wondered where birds went once the weather became really cold while growing up? After all, human beings can easily treat themselves to a warm cup of soup in their heated- houses, and not really worry about groceries as there is a supermarket nearly everywhere nowadays. Birds, on the contrary, and particularly neotropical ones, such as the so-called trans-Gulf migrants (TGMs), usually fly over the Gulf of Mexico on a non-stop route, venturing out every Spring season, and travelling long-distances from Central and South America to higher latitudes for food and breeding opportunities (@newton2011ecology). The arrival date for these birds is not a random effect though, for it has direct implications on fitness. In order words, arriving too early can increase and individual’s chances of finding a good mate and procreating, whereas arriving too late can perhaps allow an individual to find more food sources available (@smith2005arrival). Considering the importance of arrival time in the context of a changing climate, studies in the late 90s have shown variation in the arrival time of certain species, such as warblers, on a year to year basis, which was found to be linked to different environmental conditions (@smith2005arrival). However, in what specific ways these birds manage to shift their arrival and whether they actually do so remain up for a debate. 

In this project, we will explore available resources like eBird and NOAA’s National Climatic Data Center’s API in order to observe and study the effects of changes in local weather and environmental aspects linked to climate change and how these affect the migration of 10 species of TGMs. 

In undertaking these analyses, we focused on a few key questions:

•	Is there an actual variation between arrival time and wind variables along migration route for TGMs migrating to MA?

•	If so, what role can climate change potentially play in the population status of TGMs arriving in MA during the spring?

•	Is there a variation among different species? In other words, does weight affect this variation at all?

•	How does this project contribute to, challenge, or refine previous hypothesis concerning the role that climatic variables play in long-distance migration in passerine birds?

Overall, answering these questions will not only provide a better understanding of how and if individual TGMs species are able to adjust their arrival time based on a changing climate and still successfully find suitable mates and food sources, but also show evidence that climate change is an ongoing phenomenon.


## Methods

```{r,echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}

species <- c("Sphyrapicus varius","Archilochus colubris","Vermivora cyanoptera","Parkesia noveboracensis","Buteo platypterus","Contopus virens","Coccyzus erythropthalmus","Myiarchus crinitus","Catharus ustulatus","Piranga olivacea")
y <- paste0("1990",",","2019")
m <- paste0("3",",","6")

dat.l <-list()
for(s in species){
  n.obs <-  occ_data(scientificName=s,year=y,month=m,limit=0,country="US",basisOfRecord = "HUMAN_OBSERVATION",stateProvince="Massachusetts")$meta$count 
  print(n.obs)
  
  dat.l[[paste0(s)]] <- occ_data(scientificName = s,year=y,month=m,
                                 limit=n.obs,country="US",
                                 basisOfRecord = "HUMAN_OBSERVATION",
                                 stateProvince="Massachusetts")[[2]]
}
dat <- rbindlist(dat.l,fill=T)
head(dat)

saveRDS(data,"massbird.data.RDS")
library(tidyverse)
```
```{r,echo=FALSE,message=FALSE, warning=FALSE}
dat%>%
  group_by(year,species)%>%
  summarise(count=sum(individualCount,na.rm = T))%>%
  ggplot(aes(x=year,y=count,col=species))+geom_point()
```
Recorded observations of each species
```{r,echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
###NOAA Stuff
options(noaakey = "xIQmhYFPwNFiyWfXvHvzdLNfxaGypXmE")
sts <- c(
  "GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
  "GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
  "GHCND:USW00014739" #Boston
)
sta.d <- bind_rows( #bind the rows
  lapply(sts,function(x) ncdc_stations(stationid = x)$data ) #use lapply to run through stations
)%>%
  left_join(usmap_transform(.[,c("longitude","latitude")]))%>% #join transformation of lat/long for projection with usmap
  mutate(name=str_sub(name, -5,-4))%>%#simplify the name column, grab just the state
  mutate(migr.day=c(10,5,0))%>% #so we can look at wind speed 0, 5 or 10 days before arrive in boston
  separate(id,into = c("station.type","id"))%>%#need to cut station type out from station id number
  print()
weather.d <- meteo_pull_monitors(sta.d$id,date_min = "2000-01-01")
head(weather.d)


#Arrival date
mc<- dat%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)

mc.pred <- mc%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(mc%>%dplyr::select(j.day,date))

mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

#normalizing wind

weather.d <- weather.d%>%
  dplyr::mutate(year=as.integer(str_sub(date,1,4)), 
                date=as.Date(date))%>%
  group_by(year)%>% 
  dplyr::mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), 
                date2=date,
                wdir.rad=(200-abs(wdf2-200))*pi/180, 
                wvec=cos(wdir.rad)*-1*awnd 
  )%>% #store day in new column
  dplyr::select(id,year,date2,j.day,tmin,tmax,wvec)%>% 
  left_join(sta.d%>%dplyr::select(id,name,migr.day))%>% 
  dplyr::mutate(j.day=j.day+migr.day)

#Add wind/species

mc.arr.weath <- mc.arrive.date%>%
  left_join(weather.d)%>%
  left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  dplyr::select(j.day,date2,name,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
  left_join(weather.wk)
head(mc.arr.weath2)

#Mixed effect linear modeling
mc.lmer <- lmer(j.day~wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)


species.mass <- read_csv("species.mass.txt")

mc.arr.weath.mass <- mc.arr.weath%>%
  left_join(species.mass)%>%
  print()

mc.arr.weath2 <- mc.arrive.date%>%
  left_join(species.mass)%>%
  left_join(weather.wk)

head(mc.arr.weath2)

#Analysis/modeling

mc.lmer <- lmer(j.day~mass*wvec+(1|name),mc.arr.weath.mass,na.action = "na.fail")
Anova(mc.lmer)

mc.lmer2 <- lmer(j.day~mass*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2)
```

```{r,echo=FALSE,message=FALSE, warning=FALSE}
mc.arr.aic <- dredge(mc.lmer2)

mc.kb <- knitr::kable(mc.arr.aic[1:3,],caption = "Fit values for nested models of the lme model")
kable_styling(mc.kb)
```
##Results
Anova tests on one-day data, vs the two week mean data reveal that the two week mean data demonstrates a significant fixed effect of mass and wind vector. Further, the dredge table shows how a wind:mass model has an AICc value of 2202.674, as opposed to the AICc value of 2206.764 for a model relying solely on wind vector. The slope of the day of arrival vs mass model is -0.0008587, whereas the slope of the arrival day vs wind vector is 0.0010622.
```{r,echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
best.lmer <-  lmer(j.day~mass+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")

Anova(best.lmer)
```

## Discussion
The question we asked at the start of the process was whether wind vector that we would calculate based off of wind speed and the angle of the wind, and also the mass of the bird species would affect their arrival time. Through our experiment we proved that between the modeling including our calculated wind vectors a single variable and our wind vector and mass as covariates, the second model including the wind vector and mass proved to be a more accurate and statistically significant model. This connects to previous studies done in the same field as we also look at how exactly does wind and mass affect migratory bird flight, as we now have proven that wind and mass both have an effect on a migratory bird's travel. First looking at the mass's effect on migratory bird travel, we find that about 89.9% of variation in arrival time, and in a larger sense, migration pattern of different species of birds, is explained by mass, and phenology that is essentially connected to mass (@horton2019scaling). We also find in our data how largely mass affects the variation of the models through us observing the difference in statistical significance of models including single wind vector variable or wind vector and mass covariate. Additionally, looking at the comparison of the simple modeling of arrival time and  

## Author Contributions

In an effort to contribute to each other's learning as a whole, Molecool team members worked together to tackle the code, downloading and normalizing data, culminating in an organized final .Rmd document. Team member Stephanie contributed by helping crack the code, and qualitatively assessing the effects of wind and bird mass and their respective results; team member Jessica contributed to the project by doing research to put together a comprehensive introduction, the bib file, citations, and coding; team member Greg contributed to the project by halping with the code, writing the discussion, and the Rmd file; team member Julia contributed to the project by doing creating a csv file containing bird mass according to species, writing up the methods, and heling crack the code as well.

## References