---
title: "molecool_Module8"
author: "molecool"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(tidyverse)
library(ape)
library(nlme)
library(geiger)
library(caper)
library(phytools)
library(viridis)
library(MuMIn)
```
## Introduction

## Methods 
The basis of the project was to identify through experiment the relationship for force and length, and how fatigue affects the relationship between the two. We attempt to find the relationship through our experiment of measuring the maximum isometric force exerted through different angles, and measuring the force before and after fatigue of the measuring arm. We measure the maximum isometric force through what is called a DAQ pack. It contains two hangers where one attaches under something immobile like a heavy chair, and the other hook is connected to the handle that you exert your force upon by pulling in the opposite direction from where the bottom hook is. The DAQ pack is connected to the Arduino unit and it measures the amount of force that is exerted upon by pulling at the DAQ pack. We measure the varying maximum isometric force exerted through different angles by varying the angle that your forearm makes with your upper arm, the force-angle curves of the human forearm flexors, as this represents different Force-Length relationships of the arm and its torque. We measure the maximum voluntary contraction force through 12 different angles the arm can make, and the same procedure is repeated after eccentric fatigue, where we tire the forearm flexor through lowering our forearm carrying relative heavy mass from the position of the forearm and the upperarm making a 90 degree angle, and bringing the arm back up while bringing the heavy mass up with the other arm, and repeating this exercise for 3 minutes, After the fatigue, we measure the maximum isometric force of the same 12 angles. Additionally,

## Results

## Discussion

## Author Contributions

## References 